---
title: "An Introduction to the `fars` Package"
author: "Kim Aquilino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

The `fars` package provides various functions to read and summarize data from the Fatality Analysis Reporting System (FARS) at https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars. 

## fars_read

`fars_read` reads the FARS `filename` in csv format and creates a dplyr tbl_df from it. If the file or permission to it does not exist, an error is produced.

```{r fars_read}
library(fars)
test_data <- system.file("extdata", "accident_2013.csv.bz2", package = "fars")
fars_read(test_data)
```

## make_filename

`make_filename` generates a character string of the FARS filename given the year. The filename follows the format `accident_YYYY.csv.bz2` where YYYY is the 4 digit year. 

```{r make_filename}
make_filename(2014)
make_filename('2014')
make_filename(2015-1)
```

## fars_read_years

`fars_read_years` takes a vector of four digit years (coerced to integer), reads the series of FARS files and returns a dplyr tbl_df of the month and year columns for each file.

```{r fars_read_years}
fars_read_years(2013:2015)
fars_read_years(2013)
fars_read_years(c(2013,2015))
```

## fars_summarize_years

`fars_summarize_years` reads a series of FARS files and returns a frequency table of fatal crashes by month and year.

```{r fars_summarize_years}
fars_summarize_years(2013:2015)
fars_summarize_years(2013)
```

## fars_map_state

`fars_map_state` reads a FARS file and plots a map of fatal crashes.

```{r fars_map_state}
library(maps)
fars_map_state(1,2014)
fars_map_state(4,2015)
```
